---
title: "Section 2.1 Probability Basics"
output: 
    learnr::tutorial:
      progressive: true
      allow_skip: true
      id: math350_lrt_2_1_Simulations
      version: 1 # change the version when you want to invalidate past progress
runtime: shiny_prerendered
description: > 
  Probability is a way to mathematically describe random events. 
---

```{r setup, include=FALSE}
library(learnr)
library(gradethis)
library(learnrhash)
gradethis::gradethis_setup()
knitr::opts_chunk$set(echo = FALSE)
tutorial_options(
  exercise.blanks = TRUE, 
  exercise.reveal_solution = FALSE
)
```

## Introduction
These tutorials offer a place for you to practice with R code, but also a check on your understanding of the material in the course notes. Sometimes you will be asked to write sentence answers, such as in the vocabulary section below. However due to the format of these tutorials, only objects stored as R code are contained in your submission. A workaround will be to write your sentence answer and store it as an R object 

```R
answer <- "I will practice hard and do good in this class."
```

The other option is to directly print out the sentence

```R
print("This is another way to pass a sentence as R code")
```

Either way is fine. 

## Vocabulary

Define each of the terms below. 
```{r vocab, exercise = TRUE}
experiment <-  "_____"
outcome <- "_____"
sample_space  <- "_____"
event  <- "_____"
trial  <- "_____"
```


For each of the following problems, identify the sample space and the events described in set notation.  

1. Observe eye color of a group of students.

```{r vocab-yti1, exercise = TRUE}
sample_space  <- "_____"
event  <- "_____"
```

2. Number of credits a student can take.

```{r vocab-yti2, exercise = TRUE}
sample_space  <- "_____"
event  <- "_____"
```

3. Toss a coin and roll a die.

```{r vocab-yti3, exercise = TRUE}
sample_space  <- "_____"
event  <- "_____"
```

4. A soccer team is in the playoffs. The team will play three games and will either win (`w`) or lose (`l`) each game (assume ties are not allowed).

```{r vocab-yti4, exercise = TRUE}
sample_space  <- "_____"
event  <- "_____"
```


## Set operations in `R`

The following functions are used to compute intersections and unions. Each function can only take into consideration 2 vectors. 

* `union`:  Either event, or both. ($\cup$)
* `intersect`: Where do the two events overlap. ($\cap$) 
* `setdiff`: Where the two events do not overlap.

### Example

Lets revisit the deck of cards problem from above: _A single card is drawn from a standard deck of cards. Let $A$ be the event that an ace is selected, and let $B$ be the event that a heart is drawn._

First create the sample space. I am also printing the vector to ensure that what's being created is what is intended. Trust, but verify your code. 

```{r samplespace1, exercise = TRUE}
(numbers <- rep(c(2:10, "J", "Q", "K", "A"), 4))
(suits <- rep(c("H", "C", "D", "S"), each = 13))
(deck <- paste0(numbers, suits))
```

```{r samplespace1-solution}
(numbers <- rep(c(2:10, "J", "Q", "K", "A"), 4))
(suits <- rep(c("H", "C", "D", "S"), each = 13))
(deck <- paste0(numbers, suits))
```

```{r samplespace1-check}
grade_code()
```

Define event A using the elements defined in the `deck`, e.g. `"AH", "AC", "AD", "AS"`. 
```{r eventA, exercise = TRUE, exercise.blanks = TRUE}
(aces <-___) 
```

```{r eventA-solution}
(aces <- c("AH", "AC", "AD", "AS")) 
```

```{r eventA-check}
grade_this({
  if(length(.result) < length(.solution)){
    fail("It seems you're missing a card or two...")
  }
  
   if(length(unique(.result)) < length(unique(.solution))){
    fail("I spy a duplicate card.")
   }
  
  if (setequal(.result, .solution)) {
    pass("Great work!")
  }
  
  fail("Try again! Do your cards match those found in the deck? (e.g. 'AC' not 'AceClubs')")
})
```

Define event B (again using elements defined in `deck` e.g `2H`. 
```{r eventB, exercise = TRUE, exercise.blanks = TRUE}

```

```{r eventB-solution}
(hearts <- paste0(c(2:10, "J", "Q", "K", "A"), "H"))
```

```{r eventB-check}
grade_this({

  if(length(.result) < length(.solution)){
    fail("It seems you're missing a card or two...")
  }
  
  if(length(unique(.result)) < length(unique(.solution))){
    fail("I spy a duplicate card.")
  }
  
  if (setequal(.result, .solution)) {
    pass("Great work!")
  }
  
  fail("Try again! Do your cards match those found in the deck? (e.g. '2H' not 'H2')")
})
```

Use R functions to find the following statements. <span style="color:red">_In these tutorials R code chunks don't talk to each other so you have to define your vectors each time. This does NOT happen in homework._</span>

* $A\cup B$ 

```{r union, exercise = TRUE, exercise.blanks = TRUE}
aces <- ___
hearts <- ___
(aces.and.hearts <- ____(___, ____))
```

```{r union-solution}
aces <- c("AH", "AC", "AD", "AS")
hearts <- paste0(c(2:10, "J", "Q", "K", "A"), "H")
(aces.and.hearts <- union(aces, hearts))
```

```{r union-check}
grade_this({
 if(length(.result) < length(.solution)){
    fail("It seems you're missing a card or two...")
  }

  if (setequal(.result, .solution)) {
    pass("Great work!")
  }
  
  fail()
})
```

* $A\cap B$ 

```{r intersect, exercise = TRUE, exercise.blanks = TRUE}
aces <- ___
hearts <- ___
(ace.of.hearts <- ____(___, ____))
```

```{r intersect-solution}
aces <- c("AH", "AC", "AD", "AS")
hearts <- paste0(c(2:10, "J", "Q", "K", "A"), "H")
(ace.of.hearts <- intersect(aces, hearts))
```

```{r intersect-check}
grade_this({
 if(length(.result) < length(.solution)){
    fail("It seems you're missing a card or two...")
  }

  if (setequal(.result, .solution)) {
    pass("Great work!")
  }
  
  fail()
})
```

* $B^{c}$ 

```{r complement, exercise = TRUE, exercise.blanks = TRUE}
aces <-  ___
hearts <- ___
numbers <-___
suits <- ___
deck <- paste0(numbers, suits)
(no.hearts <-  ____(___, ____))
```

```{r complement-solution}
aces <- c("AH", "AC", "AD", "AS")
hearts <- paste0(c(2:10, "J", "Q", "K", "A"), "H")
numbers <- rep(c(2:10, "J", "Q", "K", "A"), 4)
suits <- rep(c("H", "C", "D", "S"), each = 13)
deck <- paste0(numbers, suits)
(no.hearts <- setdiff(deck, hearts))
```

```{r complement-check}
grade_this({
 if(length(.result) < length(.solution)){
    fail("It seems you're missing a card or two...")
  }

  if (setequal(.result, .solution)) {
    pass("Great work!")
  }
  
  fail()
})
```

## Probability rules

### You try it:

If 50 percent of the families in a certain city subscribe to the morning newspaper, 65 percent of the families subscribe to the afternoon newspaper, and 85 percent of the families subscribe to at least one of the two newspapers.

* What percentage of the families subscribe to both newspapers? (Use R as a calculator here.)

```{r ytia, exercise = TRUE}

```

::: {#ytia-hint}
Use the last bullet point in the probability axioms on page 17. 
:::

```{r ytia-check}
grade_result(pass_if(~identical(.result, .3)))
```

* What percentage of the families subscribe to only the afternoon paper? (Use R as a calculator here.)

```{r ytib, exercise = TRUE}

```

::: {#ytib-hint}
Recall $P(pm)=0.65$, represents the _entire_ circle for the afternoon paper. Find $P(PM \cap AM^{c})$. 
:::

```{r ytib-check}
grade_result(pass_if(~identical(.result, .35),))
```

* What percentage of the families don't subscribe to any paper? 

```{r ytic, exercise = TRUE}

```

::: {#ytib-hint}
Use the fact that the entire sample space has to equal 1. 
:::

```{r ytic-check}
grade_result(pass_if(~ identical(.result, .15), "Good job!"))
```

### Example

David Diez was interested in exploring the factors that contribute to an email being flagged as spam by Gmail's system. So they downloaded all their emails for a few months in 2012 and noted certain characteristics such as if it was flagged as spam (0 means no, and 1 means yes), and what size of a number it contained (none, small, or big). A _two-way table_ of emails with these two characteristics are shown below. 

```{r echo=FALSE}
em <- openintro::email
addmargins(table(em$spam, em$number, dnn=c("Spam", "Size of number"))) 
```

If you were to randomly select an email from this pool, calculate the following probabilities using `R` as a calculator: 

* It is flagged as spam: 

```{r example2b, exercise = TRUE}

```

```{r example2b-check}
grade_result(pass_if(~ identical(.result, 367/3921), "Nice!"))
```

* It has a big number: 

```{r example2c, exercise = TRUE}

```

```{r example2c-check}
grade_result(pass_if(~ identical(.result, 545/3921)))
```

* It is not flagged as spam and has a small number: 

```{r example2d, exercise = TRUE}

```

```{r example2d-check}
grade_result(pass_if(~ identical(.result, 2659/3921)))
```



### You try it
The following data table describes the sex by species breakdown for 333 observed penguins on islands in the Palmer Archipelago, Antarctica. 

```{r, echo=FALSE}
pp <- palmerpenguins::penguins
addmargins(table(pp$species, pp$sex, 
                 dnn=c("Species", "Sex")))
```

If you were to select a penguin at random from these islands, what is the estimated probability that, 

* the penguin is female

```{r penguina, exercise = TRUE}

```

```{r penguina-check}
grade_result(pass_if(~ identical(.result, 165/333)))
```


* the penguin is a Gentoo species

```{r penguinb, exercise = TRUE}

```

```{r penguinb-check}
grade_result(pass_if(~ identical(.result, 119/333)))
```


* the penguin is a male Chinstrap

```{r penguinc, exercise = TRUE}

```

```{r penguinc-check}
grade_result(pass_if(~ identical(.result, 34/333)))
```


## Submit your work
Part of your grade in this class comes from submitting tutorial work like this. 

Click the <kbd>Generate</kbd> button below to generate a hash for you assignment. This long string of numbers is a unique code for you, and this assignment. 

Paste this code into the Google Form below to submit your hash to the instructional team for scoring. You also have the opportunity to provide feedback and thoughts about the specific lesson. This is valuable information for the instructional team to know what topics were confusing for students or where extra help is needed. 

```{r context="server"}
#encoder_logic(strip_output=TRUE)
encoder_logic()
```

```{r encode, echo=FALSE}
encoder_ui(default_ui(url = "https://forms.gle/8azGaNNV73hCw19BA"))
```




